Towards a Neural Co-Processor Which Restores Dexterity After Stroke: Modeling a Proof-of-Concept

### 2021-12-05 paper/article.tex
Paper progress

TODO: summary of conclusions that will go here:
\begin{itemize}
	\item Task performance improves drastically. In all cases, performance is still improving slowly when I stop the simulation.
		\begin{itemize}
			\item 90\% on connection lesion, both with and without co-adaptation
			\item M1 50\%: 81\% recovery with and without co-adpation. The co-processor is having trouble finding a better solution than this, due to loss in brain functionality. Increasing the rate of brain co-adaptation may show a better solution.
			\item AIP 50\%: 53\% recovery w/co-adaptation so far; still running. The loss was very low to start with, and recovery depends heavily on co-adaptation, so training is slow. Without co-adaptation, the co-processor cannot learn well, because too much encoding machinery is lost.
			\item Waiting on pre-recovery result. Unclear if the co-processor will help or not.
		\end{itemize}
	\item Initially, performance improves orders of magnitude more quickly than later. 50-75\% recovery during this phase. This appears to be the co-processor learning the task structure (e.g. not to move the hand until after the hold signal is lifted).
	\item Fine tuning takes a far longer time
	\item This may have some connection to object class separation; show class sep metric side-by-side. (I have the data in hand, but haven't tried to correlate it to recovery yet).
	\item Training efficiency analysis. The majority of epochs (80\%; depends on experiment type) are EN training epochs, so most efficiency gains will be there.
	\item Can co-adapt with the brain, as it changes
	\item Compare co-adaptive case vs not. Some tasks will be impossible to accomplish without the two cooperating, if enough cortex has been lost. (waiting for new results to confirm)
	\item Show pre-recovery results; co-processor will hopefully improve beyond what recovery can achieve. (Waiting on results; running now)
\end{itemize}

@@

Abstractly, the co-processor lends itself to any closed-loop neural stimulation problem
where a relevant stimulation model can be identified. There are many lower dimensional problems
today which we address with far simpler stimulation regimes than would be necessary for
e.g. remediating the symptoms of a stroke. In some applications, such as PD symptom relief,
stimulation is often parameterized by a single off-vs-on parameter, or a small handful of
voltage parameters, where the key challenge is learning when to apply the stimulation, and
its magnitude. Our simulation demonstrates the ability of our training method to adapt to
higher dimensional problems than these, where additionally the long-running effects of
stimulation on the intended figure-of-merit must be modeled. These properties together
suggest it may be useful in lower dimensional problems where a stimulation model can be
estimated, and stimulation parameters can be tied to the target objective.

@@

Having now demonstrated our method on a higher dimensional problem, we will next demonstrate
it on a lower dimensional problem, one which is likely to transfer more readily to
an \textit{in vivo} trial. We hypothesize that this co-processor model will transition well
to applications in PD, or essential tremor (ET). In Castaño-Candamil et al. \cite{castano.pd},
we see an ML-based approach to closed-loop stimulation for ET relief. Here, a simple learning
model extracts neural markers (NMs) from a human patient's M1 cortex. It uses those to
condition DBS stimulation parameters. The stimulation controller adapts to non-stationarity
of the mapping between NMs and optimal control, which is necessary due to the sensitivity
of NMs to context (e.g. sitting or standing). The co-processor model is likely to
transition well to such an application, where it may provide additional energy savings
or reduction of stimulation side-effects. Here, it will attempt to extract neural
markers appropriate for each context, to detect transitions in context, and to apply
optimal stimulation.

>>>>

TODO: summary of conclusions that will go here:
\begin{itemize}
	\item Task performance improves drastically. In all cases, performance is still improving slowly when I stop the simulation.
		\begin{itemize}
			\item 90\% on connection lesion, both with and without co-adaptation
			\item M1 50\%: 81\% recovery with and without co-adpation. The co-processor is having trouble finding a better solution than this, due to loss in brain functionality. Increasing the rate of brain co-adaptation may show a better solution.
			\item AIP 50\%: 53\% recovery w/co-adaptation so far; still running. The loss was very low to start with, and recovery depends heavily on co-adaptation, so training is slow. Without co-adaptation, the co-processor cannot learn well, because too much encoding machinery is lost.
			\item Waiting on pre-recovery result. Unclear if the co-processor will help or not.
		\end{itemize}
	\item Initially, performance improves orders of magnitude more quickly than later. 50-75\% recovery during this phase. This appears to be the co-processor learning the task structure (e.g. not to move the hand until after the hold signal is lifted).
	\item Fine tuning takes a far longer time
	\item This may have some connection to object class separation; show class sep metric side-by-side. (I have the data in hand, but haven't tried to correlate it to recovery yet).
	\item Training efficiency analysis. The majority of epochs (80\%; depends on experiment type) are EN training epochs, so most efficiency gains will be there.
	\item Can co-adapt with the brain, as it changes
	\item Compare co-adaptive case vs not. Some tasks will be impossible to accomplish without the two cooperating, if enough cortex has been lost. (waiting for new results to confirm)
	\item Show pre-recovery results; co-processor will hopefully improve beyond what recovery can achieve. Initial result: the co-adaptive case found a better solution. The co-processor couldn't improve yet on the pre-recovery case. The solution it found was (basically) to ignore the inputs.
\end{itemize}

@@

Abstractly, the co-processor lends itself to any closed-loop neural stimulation problem
where a relevant stimulation model can be identified. There are many lower dimensional problems
today which we address with far simpler stimulation regimes than would be necessary for
e.g. remediating the symptoms of a stroke. In some applications, such as PD symptom relief,
stimulation is often parameterized by a single off-vs-on parameter, or a small handful of
parameters, where the key challenge is learning when to apply the stimulation, and
its shape and power. Our simulation demonstrates the ability of our training method to adapt to
higher dimensional problems than these, where additionally the long-running effects of
stimulation on the intended figure-of-merit must be modeled. These properties together
suggest it may be useful in lower dimensional problems where a stimulation model can be
estimated, and stimulation parameters can be tied to the target objective.

@@

Having now demonstrated our method on a higher dimensional problem, we will next demonstrate
it on a lower dimensional problem, one which is likely to transfer more readily to
an \textit{in vivo} trial. We hypothesize that this co-processor model will transition well
to applications in PD, or essential tremor (ET). In Castaño-Candamil et al. \cite{castano.pd},
we see an ML-based approach to closed-loop stimulation for ET relief. Here, a simple learning
model extracts neural markers (NMs) from a human patient's M1 cortex. It uses those to
condition DBS stimulation parameters. The stimulation controller adapts to non-stationarity
of the mapping between NMs and optimal control, which is necessary due to the sensitivity
of NMs to context (e.g. sitting or standing). The co-processor model is likely to
transition well to such an application, where it may provide additional energy savings
or reduction of stimulation side-effects. In that case, it will attempt to extract neural
markers appropriate for each context, to detect transitions in context, and to apply
optimal stimulation.
